An investigation into using scagnostics to find interesting projections of high-dimensional data

### 2018-10-24 scagnostics.Rmd
Added short example of index parameter dependence.

mineIndex <- function(mineIndex){
  function(mat){
    return(mine(mat[,1], mat[,2])[[mineIndex]])
  }
}
```

@@

>>>>

mineIndex <- function(mineIndex){
  function(mat){
    return(mine(mat[,1], mat[,2])[[mineIndex]])
  }
}

mineIndexAlpha <- function(mineIndex, alpha){
  function(mat){
    return(mine(mat[,1], mat[,2], alpha=alpha)[[mineIndex]])
  }
}
```

@@

### Illustrative example
A nice example can be constructed based on the RANDU dataset available in R. We use the MIC index and compare different parametrisations of the maximum binning by $\alpha$, where $B(n) = n^{\alpha}$. We compare in Figure \ref{fig:randuEx} the final view, index value and computing time obtained when optimising the MIC index with values $\alpha = 0.6, 0.7, 0.8$. Despite the low dimensionality, the small quint angle of the structure means that the identification of the maximum depends strongly on the starting conditions. Empirically we have found that using $\alpha = 0.7$ is most efficient in identifying the structure, while the default value $\alpha = 0.6$ is most likely to miss the structure given equal starting conditions.

\textcolor{red}{Possible interpretation: 0.6 not allowing enough bins to accurately capture the structure, 0.7 good compromise, 0.8 becomes overly sensitive to noise}

```{r randuEx, results="hide", fig.cap="Final view when optimising the MIC index in the RANDU dataset. Each plot lists the selected value of $\\alpha$, the time required by the optimisation in seconds and the MIC index value."}
if(!file.exists("cache/randuEx.rda")){
  pL <- list()
  i <- 1
  for (alpha in c( 0.6, 0.7, 0.8)){
    set.seed(556677)
    tic.clearlog()
    tic() #start timer
    tP <- save_history(randu, guided_tour(mineIndexAlpha("MIC", alpha)),max_bases = 100000)
    toc(log=TRUE,quiet=TRUE)
    optT <- unlist(tic.log(format=FALSE))["toc.elapsed"]-unlist(tic.log(format=FALSE))["tic.elapsed"]
    imax <- length(as.list(tP))
    fV <- as.tibble(as.matrix(randu) %*% as.list(tP)[[imax]])
    colnames(fV) <- c("PP1", "PP2")
    idxV <- mine(fV$PP1, fV$PP2, alpha = alpha)[["MIC"]]
    pL[[i]] <- ggplot(fV, aes(PP1,PP2)) + geom_point() + theme(aspect.ratio=1) + 
      ggtitle(paste(alpha, toString(format(optT, digits=2)), toString(format(idxV, digits=2)), sep = ", "))
    i <- i+1
  }
  save(pL, file = "cache/randuEx.rda")
} else {
  load("cache/randuEx.rda")
}
grid.arrange(pL[[1]], pL[[2]], pL[[3]], ncol=3)
```
